### 2011-06-13 thesis/novelty.tex
Fixes here and there.

\begin{eqnarray*}
 P_B(c=0 | x)         &<& P_B(c=1 | x) \\
\Phi_B(c=0 \land x)   &<& \Phi_B(x=1 \land x) \\
\Phi_G(x) \phi_t(c=0) &<& \Phi_{U_a}(x) \phi_t(c=1) \\
P_G(x) Z_G \phi_t(c=0) &<& P_{U_a}(x) Z_{U_a} \phi_t(c=1) \\
\frac{P_G(x)}{P_{U_a}(x)} &<& \frac{Z_{U_a}}{Z_G}
\end{eqnarray*}

@@

This equation is not very practical due to the need of calculation of
normalization factors $Z_G$ and $Z_{U_a}$. The presence of this factors
not only turns computation more expensive, but also requires that a full model
is known for all the factors disabling the ability to incorporate probabilistic
sensing output from complex detectors such as \glspl{SVM}. Ultimately, it also
removes capability to mix complex thresholds on several variables that would
allow to do probabilistic inference on the novelty of multiple variables at once.

@@

\subfloat[Training Data]{
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [blue] {}; \& \node (e) [yellow] {}; \\
    \node (b) [yellow] {}; \& \node (f) [blue] {}; \\
    \node (c) [blue] {}; \& \node (d) [yellow] {}; \\
  };
  \draw [-] (a) -- (e) -- (f) -- (d) -- (c) -- (b) -- (f);
\end{tikzpicture}
\quad
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
    \node (empty) [blue,fill=none,draw=none] {}; \& \\
    \node (c) [blue] {}; \& \node (d) [yellow] {}; \\
  };
  \draw [-] (c) -- (a) -- (d) -- (e);
\end{tikzpicture}
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
    \node (b) [blue] {}; \& \node (f) [yellow] {}; \\
    \node (c) [yellow] {}; \& \node (d) [blue] {}; \\
  };
  \draw [-] (a) -- (e) -- (f) -- (d) -- (c) -- (b) -- (a);
\end{tikzpicture}
\quad
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [blue] {}; \& \node (e) [yellow] {}; \\
    \node (b) [yellow] {}; \& \node (f) [blue] {}; \\
    \node (c) [blue] {}; \& \node (d) [blue] {}; \\
  };
  \draw [-] (c) -- (b) -- (a) -- (e) -- (f) -- (b);
  \draw [-] (b) -- (d);
\end{tikzpicture}
\quad
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
    \node (b) [blue] {}; \& \node (f) [yellow] {}; \\
    \node (c) [yellow] {}; \& \node (d) [blue] {}; \\
  };
  \draw [-] (e) -- (a) -- (b) -- (c) -- (d);
  \draw [-] (b) -- (f);
\end{tikzpicture}

@@

Note that with this transformation, the following potential equivalences hold:
\begin{eqnarray*}
\Phi_{G'}(x, c=0) &=& \Phi_{G}(x | a \in \{A_1, \dots, A_K\})\\
\Phi_{G'}(x, c=1) &=& \Phi_{G}(x) 
\end{eqnarray*}

Although those potentials allow to determine novelty of $a$, they require
knowledge of all the possible classes, which is not available.
Nonetheless access to unlabelled data or other informations or assumptions on
the unconditional distribution allow to model or approximate unconditional
distributions of $a$. For that there is interest in trying to model the set
of states $\{A^*_1, \dots, A^*_N\}$ with a set of states $\{A^{\dag}_1, \dotsc, A^{\dag}_M\}$.

Moreover assuming that the set of states $A^\dag$ is able to perfectly
characterize the distribution over any graph $G$,
for a specific graph the potentials with states $A^\dag$ are a scale-factor
away from the potentials with states $A$. And that scale-factor is dependent on
the graph structure and factors of both $G$ and $G^\dag$:

@@

Note that this ratio, although seems similar to the one obtained in
\autoref{eq:structure-ratio} it has a special property. Here both $G$ and
$G^\dag$ model the same probability, so it is expected that both scale they
normalization factor according to some relation on how a given factor differs
from $G$ to $G^\dag$.\footnote{On \autoref{eq:structure-ratio} the ratio is
between two different distributions: a conditional and an unconditional distribution.}

If on any graph structure the marginal probabilities of all variables except $a$
is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$
then $\sum_{A^\dag}\phi^\dag_i(a^\dag, \cdot) = s_i \sum_A \phi_i(a, \cdot)$.
It can be proved that since the distribution over any graph structure has to be
equal, then the same holds for a graph structure $P_{G_i}$ composed by a single
factor $\phi_i$:

\begin{eqnarray*}
P^\dag_{G_i}(x) &=& P_{G_i}(x) \\
\frac{\phi_i(x)}{Z^\dag_{G_i}}  &=& \frac{s_i \phi_i(x)}{Z_{G_i}} \\
s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}} \quad \text{(constant)}
\end{eqnarray*}

@@

If factor $\phi_i$ appears $k_i$ times on a graph structure then:
\begin{eqnarray}
\Phi^\dag_G(x) \prod {s_i}^{k_i} = \Phi_G(x)
\end{eqnarray}

Alternative, if factors $\phi^\ddag_i(a^\dag,x) = \phi^\dag_i(a^\dag,x)/s_i$ are
introduced then:
\begin{eqnarray}
\Phi^\ddag_G(x) = \Phi_G(x)
\end{eqnarray}

Having shown how to replace a variable $a$ for an equivalent $a^\dag$, and
respectively change the factors dependent on $a$ from $\phi_i(a,\cdot)$ to
$\phi^\ddag_i(a^\dag,\cdot)$ such that the potentials on the original
graph are kept constant it becomes possible to go back to the graph $G'$ and
replace the states $\{A^*_1,\dotsc,A^*_N\}$ for the equivalents $A^\ddag$
and change the factors associated with $a'$ for:

@@

It is now possible to perform a decision on the novelty of $a$ with the condition:
\begin{equation}
2 \Phi_{G'}(x | c = 0) < \Phi_{G'}(x | c= 1)
\end{equation}

>>>>

\begin{eqnarray*}
 \frac{P_G(x)}{P_{U_a}(x)}&<& t_A \\
 \frac{\Phi_G(x)}{\Phi_{U_a}(x)}    &<& t_A \frac{Z_G}{Z_{U_a}} \\
 \frac{\Phi_B(x, c=0)}{\Phi_{B}(x, c=1)}    &<& t_A \frac{Z_G}{Z_{U_a}} \\
\end{eqnarray*}

@@

This equation is not very practical due to the need of calculation of
normalization factors $Z_G$ and $Z_{U_a}$. The presence of this factors
not only turns computation more expensive, but also requires that a full model
is known for all the factors disabling the ability to incorporate probabilistic
sensing output from complex detectors such as \glspl{SVM}.

@@

\subfloat[Training Data]{
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [blue] {}; \& \node (e) [yellow] {}; \\
    \node (b) [yellow] {}; \& \node (f) [blue] {}; \\
    \node (c) [blue] {}; \& \node (d) [yellow] {}; \\
  };
  \draw [-] (a) -- (e) -- (f) -- (d) -- (c) -- (b) -- (f);
\end{tikzpicture}
\qquad
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
    \node (empty) [blue,fill=none,draw=none] {}; \& \\
    \node (c) [blue] {}; \& \node (d) [yellow] {}; \\
  };
  \draw [-] (c) -- (a) -- (d) -- (e);
\end{tikzpicture}
\qquad
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
    \node (b) [blue] {}; \& \node (f) [yellow] {}; \\
    \node (c) [yellow] {}; \& \node (d) [blue] {}; \\
  };
  \draw [-] (a) -- (e) -- (f) -- (d) -- (c) -- (b) -- (a);
\end{tikzpicture}
\qquad
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [blue] {}; \& \node (e) [yellow] {}; \\
    \node (b) [yellow] {}; \& \node (f) [blue] {}; \\
    \node (c) [blue] {}; \& \node (d) [blue] {}; \\
  };
  \draw [-] (c) -- (b) -- (a) -- (e) -- (f) -- (b);
  \draw [-] (b) -- (d);
\end{tikzpicture}
\qquad
\begin{tikzpicture}
  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
    \node (b) [blue] {}; \& \node (f) [yellow] {}; \\
    \node (c) [yellow] {}; \& \node (d) [blue] {}; \\
  };
  \draw [-] (e) -- (a) -- (b) -- (c) -- (d);
  \draw [-] (b) -- (f);
\end{tikzpicture}

@@

Note that with this transformation, the following potential equivalences hold:
\begin{eqnarray*}
\Phi_{G'}(x, c=0) &=& \Phi_{G}(a \in \{A_1, \dots, A_K\}, x)\\
\Phi_{G'}(x, c=1) &=& \Phi_{G}(x) 
\end{eqnarray*}

Although those potentials allow to determine novelty of $a$, they require
knowledge of all the possible classes, which is not available.
But by having access to unlabelled data, other informations or assumptions on
the unconditional distribution it is possible to model the unconditional
distributions of $x$. For that there is interest in trying to model the real set
of states $\{A^*_1, \dots, A^*_N\}$ with a set of states $\{A^{\dag}_1, \dotsc, A^{\dag}_M\}$
learned or assumed from information on the unconditional distribution.

When changing a variable states from $A^*$ into $A^\dag$ it also becomes necessary to redefine
the factors dependent on $a$ to accommodate the new states $A^\dag$.
Moreover assuming that the change of variable states $A^\dag$ is able to perfectly
characterize the distribution over any graph structure,
for a specific graph $G$ the potentials $\Phi^\dag_G(x)$ with states $A^\dag$ are a
scale\hyp{}factor away from the potentials $\Phi_G(x)$ with states $A$.
And that scale-factor is dependent on the graph structure and factors of both $G$ and $G^\dag$:

@@

Note that this ratio, although seems similar to the one obtained in
\autoref{eq:structure-ratio} it has a special property. Here both $G$ and
$G^\dag$ model the same distribution, so it is expected that both scale they
normalization factor according to some relation on how a given factor differs
from $G$ to $G^\dag$.\footnote{On \autoref{eq:structure-ratio} the ratio is
between two different distributions: a conditional and an unconditional distribution.}

If on any graph structure the marginal probabilities of all variables except $a$
is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$
then it is possible to scale each individual factor $\phi_i^\dag$ with a constant $1/s_i$
such that the potentials with $A$ will be equal to the potentials with $A^\dag$.

Since the distribution over any graph structure has to be
equal, then the same holds for a graph structure $P_{G_i}$ composed by a single
factor $\phi_i$:

\begin{eqnarray*}
P^\dag_{G_i}(x) &=& P_{G_i}(x) \\
\frac{\sum_{a^\dag \in A^\dag}\phi^\dag_i(a^\dag, x)}{Z^\dag_{G_i}}  &=& \frac{\sum_{a \in A}\phi_i(x)}{Z_{G_i}} \\
s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}} \quad \text{(constant)}
\end{eqnarray*}

@@

If factor $\phi_i$ appears $k_i$ times on a graph structure then:
\begin{eqnarray}
\frac{\Phi^\dag_G(x)}{\Phi_G(x)} =  \prod {s_i}^{k_i}
\end{eqnarray}

Alternative, if factors $\phi^\ddag_i(a^\dag,x) = \phi^\dag_i(a^\dag,x) s_i$ are
introduced then:
\begin{eqnarray}
\Phi^\ddag_G(x) = \Phi_G(x)
\end{eqnarray}

Having shown how to replace a variable $a$ for an equivalent $a^\dag$, and
respectively change the factors dependent on $a$ from $\phi_i(a,\cdot)$ to
$\phi^\ddag_i(a^\dag,\cdot)$ such that the potentials are kept equal equal
it becomes possible to go back to the graph $G'$ and
replace the states $\{A^*_1,\dotsc,A^*_N\}$ for the equivalents $A^\ddag$
and change the factors associated with $a'$ for:

@@

It is now possible to perform a decision on the novelty of $a$ with the condition:
\begin{equation}
2 \Phi_{G'}(x, c = 0) < \Phi_{G'}(x, c= 1)
\end{equation}
